Title here

### 2017-01-03 portfolio/text/ms.Rmd
Edits

Individuals who rely on the extraction of natural resources for their
livelihood --- fishers, farmers, miners, and loggers, among others --- face
high levels of income variability driven by a mix of environmental, biological,
and economic variability. A key component to managing these industries is
identifying how regulatory actions and individual behaviour affect income
variability and by extension the economic stability and sustainable use of
natural resources. For commercial fisheries, recent work has identified that
communities and vessels that fish a greater diversity of species have less
revenue variability than those that fish a lower species diversity. However, it
is unclear if these benefits extend to the level of the individual and how
year-to-year changes in diversification affect revenue and revenue variability.
Here we identify three axes by which fishers in Alaska, one of the world's most
productive fishing regions, can diversify fishing activities and investigate
how individual actions along each axis affect revenue and revenue variability.
We show that fishing permits with higher species diversity reduces individual
revenue variability and fishing an additional permit is associated with higher
revenue and lower variability. However, increasing species diversity within the
constraints of a permit set has a fishery-dependent effect on revenue and is
usually (`r g2_dens`% probability) associated with increased revenue
uncertainty the following year. Our results demonstrate that the most effective
option for individuals to decrease revenue variability is to purchase permits
to participate in additional or more diverse fisheries. However, this option is
costly and often limited by regulations such as limited entry and catch share
programs. With increasing environmental variability in a changing climate, it
will be particularly important that individuals who rely on natural resources
for their livelihood have effective strategies to reduce income variability.

@@

Using a unique dataset collected over 30 years (1985--2014) on individuals,
rather than fleets or vessels, we examine how revenue and revenue variability of
individuals is affected by diversification. By tracking individuals we can track
people that switch fishery permits over time or use multiple vessels. Similarly,
we can disentangle fishers that share the same vessel. We develop and apply
a hierarchical Bayesian model to understand the effect of fisheries catch
species diversity on individual revenue and revenue variability that builds on
results from vessel [@kasperski2013], fleet [@sethi2012b], and community-level
[@sethi2014] variability.

We develop a variance-function regression model [@davidian1987; @hall1989], to
simultaneously models changes in the mean and variance of individual revenue as
a function of covariates. We simultaneously estimate the effect of species
diversity (the inverse of Simpson's diversity index [@simpson1949; @jost2006])
on individual revenue and revenue variability, and estimate effects both within
and across fishing strategies while controlling for fishing effort and latent
time effects (e.g. changing environmental conditions). To our knowledge, this is
the first application of this type of model in any natural resource field.

@@

Individual revenue variability decreases exponentially with higher mean fished
species diversity across permit sets (Fig. \ref{fig:across}). Controlling for
effort, strategies with one unit higher mean effective species diversity had on
average `r h1[[2]]`% (`r h1[[1]]`--`r h1[[3]]`% 95% CI) lower variability. This
pattern looks qualitatively similar with or without individuals targeting
salmon. As examples, variability was highest for the herring roe gillnet and
king crab fisheries. Most strategies with the lowest estimated revenue
variability combined halibut with a salmon permit and had high levels of fished
species diversity (Fig. \ref{fig:across}).

@@

In contrast to the clear across-strategy effects of species diversity on revenue
and revenue variability, the effect of changing fished species diversity for an
individual *within* the constraints of a given permit set is more variable (Fig.
\ref{fig:within}). After accounting for fishing effort, specializing tends to be
associated with increased revenue but also increased variability (Fig.
\ref{fig:within}; Figs S\ref{fig:fe}, S\ref{fig:re}, S\ref{fig:noeffort}). On
average, there is a ~0.5% increase in the year-to-year revenue ratio per 1%
decrease in species diversity, and across strategies, `r g2_dens`% of the
probability density supports a positive relationship between specialization and
revenue.

However, changing (increasing or decreasing) an individual's fished species
diversity within the constraints of a permit set tends to be associated with
greater expected variability or uncertainty in year-to-year revenue (Fig.
\ref{fig:within} strategies above zero; Figs. S\ref{fig:fe}, S\ref{fig:re},
S\ref{fig:mirror}). This effect was strongest for individuals with a halibut
permit or halibut and sablefish permits. Individuals with these strategies
experienced a ~2% increase in variability for a 1% change in the year-to-year
ratio of species diversity (Fig. \ref{fig:within}). Becoming more generalized in
species diversity also leads to increased revenue variability (Fig.
S\ref{fig:mirror}A), suggesting that the act of changing strategies is
a contributor to increased variability of revenue.

@@

Alaskan halibut and sablefish fisheries were converted to individual fishery
quota (IFQ) management in 1994, which potentially confounds our estimates with
a major regulatory change. Therefore, we refit our models allowing all
strategy-level effects to vary separately before and after the introduction of
IFQs for strategies involving halibut and sablefish permits. The introduction
of IFQs reduced variability by ~25% (TODO XX--XX 95% CI) for individuals with
only halibut or sablefish permits (Fig. S\ref{fig:ifq}A). Furthermore, for most
strategies, the association between changing species diversity and increased
revenue variability became stronger after the introduction of IFQs (Fig.
S\ref{fig:ifq}B). This was especially true for individuals fishing only halibut
or sablefish and less so for individuals combining halibut or sablefish with
a salmon permit (Fig. S\ref{fig:ifq}B).

@@

Our results across fishing strategies are in line with previous findings at the
community and vessel levels [@kasperski2013; @sethi2012b; @sethi2014], but our
results with respect to individual year-to-year behaviour provide new insights.
Within strategies, specializing tended to be associated with increased revenue.
This could be for a number of reasons. First, specialization can increase
fishing efficiency [@smith1986]. Efficiency gains from specialization can be
seen in other natural resource sectors. For example, rice farmers in Ghana can
increase their efficiency by devoting more land and labor to rice production
compared to other crops [@abdulai2000]. Second, given the availability of "cash
crops", specializing can be highly profitable. For instance, when cocoa prices
have been high, farmers in Nigeria have been highly specialized on the crop
[@mustapha1999] and specializing on valuable tobacco farming can be the optimal
strategy regardless of alternative crops [@chavez2014]. In our analyses,
fishers with the opportunity to fish highly valuable sockeye salmon benefited
the most from specializing. However, specializing can have longterm risks if
market, biological, or environmental conditions change. For example, U.S.
farmers specialized on wheat, corn, or soybeans were hit harder than
diversified farmers in the early 1980s when a financial crisis hit
[@ikerd1990]. Third, a lack of specialization can happen as a consequence of
non-ideal circumstances such as poor availability of a more valuable fish stock
in a given year. For example, downward economic shocks have been associated
with increased diversity for households relying on forests for income
[@debela2012].

@@

Our analysis provides a synthetic view of the effects of species diversity on
individuals over 30 years in one of the most productive fishing regions in the
world. However, this large temporal and geographical scope potentially obscures
time or regional-variation in the consequences of diversification on revenue
and revenue variability. We expect future research examining diversification
over finer scales will uncover additional relationships. Furthermore, in
addressing the overall pattern, we have averaged out several major discrete
events, such as the Exxon Valdez oil spill in 1989, that had significant
consequences for the ecological and social systems in Alaska [@paine1996]. In
addition to revenue or income variability, it is important to understand the
benefits of diversification in the context of rare and extreme disturbances
which occur in natural systems worldwide [REF].

@@

>>>>

Individuals who rely on the extraction of natural resources for their
livelihood --- fishers, farmers, miners, and loggers, among others --- face
high levels of income variability driven by a mix of environmental, biological,
and economic variability. A key component to managing these industries is
identifying how regulatory actions and individual behaviour affect income
variability and by extension the economic stability and sustainable use of
natural resources. For commercial fisheries, recent work has identified that
communities and vessels that fish a greater diversity of species have less
revenue variability than those that fish a lower diversity. However, it
is unclear if these benefits extend to the level of the individual and how
year-to-year changes in diversification affect revenue and revenue variability.
Here we identify three axes by which fishers in Alaska, one of the world's most
productive fishing regions, can diversify fishing activities.
We show that fishing permits with higher species diversity reduces individual
revenue variability and fishing an additional permit is associated with higher
revenue and lower variability. However, increasing species diversity within the
constraints of a permit set has a fishery-dependent effect on revenue and is
usually (`r g2_dens`% probability) associated with increased revenue
uncertainty the following year. Our results demonstrate that the most effective
option for individuals to decrease revenue variability is to purchase permits
to participate in additional or more diverse fisheries. However, this option is
costly and often limited by regulations such as limited entry and catch share
programs. With increasing environmental variability in a changing climate, it
will be particularly important that individuals who rely on natural resources
for their livelihood have effective strategies to reduce income variability.

@@

Using a unique dataset collected over 30 years on individuals,
rather than fleets or vessels, we examine how revenue and revenue variability of
individuals is affected by diversification. By tracking individuals we can track
people that switch fishery permits over time or use multiple vessels. Similarly,
we can disentangle fishers that share the same vessel. We develop and apply
a hierarchical Bayesian model to understand the effect of fisheries catch
species diversity on individual revenue and revenue variability that builds on
results from vessel [@kasperski2013], fleet [@sethi2012b], and community-level
[@sethi2014] variability.
We develop a variance-function regression model [@davidian1987; @hall1989], to
simultaneously models changes in the mean and variance of individual revenue as
a function of covariates. We simultaneously estimate the effect of species
diversity (the inverse of Simpson's diversity index [@simpson1949; @jost2006])
on individual revenue and revenue variability, and estimate effects both within
and across fishing strategies while controlling for fishing effort and latent
time effects (e.g. changing environmental conditions). To our knowledge, this is
the first application of this type of model in any natural resource field.

@@

Individual revenue variability decreases exponentially with higher mean fished
species diversity across permit sets (Fig. \ref{fig:across}). Controlling for
effort, strategies with one unit higher mean species diversity had on
average `r h1[[2]]`% (`r h1[[1]]`--`r h1[[3]]`% 95% CI) lower variability. This
pattern looks qualitatively similar with or without individuals targeting
salmon. As examples, variability was highest for the herring roe gillnet and
king crab fisheries. Most strategies with the lowest estimated revenue
variability combined halibut with a salmon permit and had high levels of fished
species diversity (Fig. \ref{fig:across}).

@@

In contrast to the clear across-strategy effects of species diversity on revenue
and revenue variability, the effect of changing fished species diversity for an
individual *within* the constraints of a given permit set is more variable (Fig.
\ref{fig:within}). After accounting for fishing effort, specializing tends to be
associated with increased revenue but also increased variability (Fig.
\ref{fig:within}; Figs S\ref{fig:fe}, S\ref{fig:re}, S\ref{fig:noeffort}). On
average, there is a ~0.5% increase in the year-to-year revenue ratio per 1%
decrease in species diversity, and across strategies weighted equally, `r g2_dens`% of the
probability density supports a positive relationship between specialization and
revenue.

Changing (increasing or decreasing) an individual's fished species
diversity within the constraints of a permit set tends to be associated with
greater expected variability or uncertainty in year-to-year revenue (Fig.
\ref{fig:within} strategies above zero; Figs. S\ref{fig:fe}, S\ref{fig:re},
S\ref{fig:mirror}). This effect was strongest for individuals with a halibut
permit or halibut and sablefish permits. Individuals with these strategies
experienced a ~2% increase in variability for a 1% change in the year-to-year
ratio of species diversity (Fig. \ref{fig:within}). Becoming more generalized in
species diversity also leads to increased revenue variability (Fig.
S\ref{fig:mirror}A), suggesting that the act of changing strategies is
a contributor to increased variability of revenue.

@@

Alaskan halibut and sablefish fisheries were converted to individual fishery
quota (IFQ) management in 1994, which potentially confounds our estimates with
a major regulatory change. Therefore, we refit our models allowing all
strategy-level effects involving halibut and sablefish permits
to vary separately before and after the introduction of IFQs. The introduction
of IFQs reduced variability by ~25% (TODO XX--XX 95% CI) for individuals with
only halibut or sablefish permits (Fig. S\ref{fig:ifq}A). Furthermore, for most
strategies, the association between changing species diversity and increased
revenue variability became stronger after the introduction of IFQs (Fig.
S\ref{fig:ifq}B). This was especially true for individuals fishing only halibut
or sablefish and less so for individuals combining halibut or sablefish with
a salmon permit (Fig. S\ref{fig:ifq}B).

@@

Our results across fishing strategies are in line with previous findings at the
community and vessel levels [@kasperski2013; @sethi2012b; @sethi2014], but our
results with respect to individual year-to-year behaviour provide new insights.
Within strategies, specializing tended to be associated with increased revenue.
This could be for a number of reasons. First, specialization can increase
fishing efficiency [@smith1986]. Efficiency gains from specialization can be
seen in other natural resource sectors. For example, rice farmers in Ghana can
increase their efficiency by devoting more land and labor to rice production
compared to other crops [@abdulai2000]. Second, given the availability of
valuable resources, specializing can be highly profitable. For instance, when cocoa prices
have been high, Nigerian farmers have tended to specialize on the crop
[@mustapha1999] and specializing on valuable tobacco farming can be the optimal
strategy regardless of alternative crops [@chavez2014]. In our analyses,
fishers with the opportunity to fish highly valuable sockeye salmon benefited
the most from specializing. However, specializing can have longterm risks if
market, biological, or environmental conditions change. For example, U.S.
farmers specialized on wheat, corn, or soybeans were hit harder than
diversified farmers in the early 1980s when a financial crisis hit
[@ikerd1990]. Third, generalization --- a lack of specialization ---
can happen as a consequence of
non-ideal circumstances such as poor availability of a more valuable fish stock
in a given year. For example, downward economic shocks have been associated
with increased diversity for households relying on forests for income
[@debela2012].

@@

Our analysis provides a synthetic view of the effects of species diversity on
individuals over 30 years in one of the most productive fishing regions in the
world. However, this large temporal and geographical scope potentially obscures
time or regional-variation in the consequences of diversification on revenue
and revenue variability. We expect future research examining diversification
over finer scales will uncover additional interesting relationships. Furthermore, in
addressing the overall pattern, we have averaged out several major discrete
events, such as the Exxon Valdez oil spill in 1989, that had significant
consequences for the ecological and social systems in Alaska [@paine1996]. In
addition to revenue or income variability, it is important to understand the
benefits of diversification in the context of rare and extreme disturbances
which occur in natural systems worldwide [REF].

@@

# Acknowledgments

TODO. National Center for Ecological Analysis and Synthesis (NCEAS) working
group *Applying portfolio effects to the Gulf of Alaska ecosystem: Did
multi-scale diversity buffer against the Exxon Valdez oil spill?*. Other
participants... Smith...
